Litrepl: Literate Paper Processor Promoting Transparency More Than Reproducibility

### 2024-12-30 paper/paper.tex
Update paper

In Litrepl, these details are hardcoded for several prominent interpreter
families, which we refer to as \textbf{interpreter classes}. At the time of
writing, Litrepl supports three classes: \textbf{python}, \textbf{sh}, and
\textbf{ai}. Using command line arguments, users can configure how to map code
section labels to the correct class and specify which interpreter command to
execute to start a session for each class.

@@

\begin{figure*}[hbt!]
  \centering
  \begin{minipage}{\textwidth}
  %lresult
  \begin{verbatim}
  \begin{result}
  ... some output ...
  [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
  \end{result}
  \end{verbatim}
  %lnoresult
  \end{minipage}
  \caption{Example output in a verbatim format displayed above two columns.}
  \label{fig:verbatim_example}
\end{figure*}

To initiate interpreter sessions, Litrepl allocates resources as depicted in
Figure \ref{fig:resource-allocation}. The process begins by creating an
auxiliary directory. The name of this directory is derived from the combination
of the current interpreter class name, the Litrepl working directory, and a
system-designated folder for temporary files. Subsequently, two POSIX
pipes—\texttt{in.pipe} and \texttt{out.pipe} are established. The configured
interpreter is then launched with these pipes attached, and the interpreter's
process identifier is recorded in a dedicated file.

The initialization routine incorporates additional open calls on the pipes to
prevent premature termination of the interpreter due to detachments. This
ensures the interpreter remains active and responsive.

When an evaluation request is received, Litrepl calculates the name of the
anticipated response storage file and forwards the request to the interpreter.
It concurrently initiates a response reader process. The response storage is
secured using soft POSIX file locks, enabling the main process to ascertain when
the response is ready. In instances where the response exceeds the configured
time to completion, Litrepl provides a partial result. This partial result can
be updated with complete details in subsequent calls. The format for a partial
result section appears as follows:

@@

\begin{figure*}[hbt!]
  \centering
  \begin{minipage}{\textwidth}
  \begin{verbatim}
  document       ::= (code | result | ignore | text)*
  code           ::= (code-normal | code-comment)
  result         ::= (result-normal | result-comment)
  code-normal    ::= "\begin{MARKER}" text "\end{MARKER}"
  code-comment   ::= "% MARKER" text "% noMARKER"
  result-normal  ::= "\begin{result}" text "\end{result}"
  result-comment ::= "% result" text "% noresult"
  ignore         ::= "% ignore" text "% noignore"
  text           ::= ...
  \end{verbatim}
  \end{minipage}
  \caption{An illustrative grammar template for LaTeX documents where marker
  serves as a parameter configured via command-line arguments for each supported
  interpreter class.}
  \label{fig:document_representation}
\end{figure*}

>>>>

In Litrepl, these details are hardcoded for several prominent interpreter
families, which we refer to as \textit{interpreter classes}. At the time of
writing, Litrepl supports three such classes: \textbf{python}, \textbf{sh}, and
\textbf{ai}. Using command line arguments, users can configure how to map code
section labels to the correct class and specify which interpreter command to
execute to start a session for each class.

@@

\begin{figure*}[hbt!]
  \centering
  \begin{minipage}{\textwidth}
  %lresult
  \begin{verbatim}
  \begin{result}
  ... some output ...
  [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
  \end{result}
  \end{verbatim}
  %lnoresult
  \end{minipage}
  \caption{Example output in a verbatim format displayed above two columns.}
  \label{fig:partial-result}
\end{figure*}

\begin{comment}
%lai
Lets rewrite the "Session management" section. Please consider the following
points, check the grammar and combine them into a linked narrative. Please
generate it using Latex markup. Wrap your own comments in Latex line comments.
Please don't generate figures. Just text.

----

1. As mentioned earlier, the necessary side-effect of code section evaluation,
which makes read-eval loop possible - is the the interpreter session left
running in the background. The resources required to manage this mode of
operation are shown on the Figure \ref{fig:resource-allocation}. All of them are
visible as files residing in some auxiliary directory.

2. If not set via arguments or environment, the path to this directory
is calculated out of interpreter class name, current working directory and the
operating system temporary folder name.

3. The code to be evaluated determines (via hashing) the name of the response file,
containing interpreter response for this code.

4. Also, aux. dir contains a couple of pipes providing an access to the
interpreter standard input and ouput streams.

5. Finally, the PID of the interpreter process is also stored in a file.

6. When evaluating, Litrepl forks a response reader which keeps it soft-locked
until the interpreter signals the completion by answering to the echo probe. If
the response takes longer than configured, Litrepl outputs a pratial response
result which can be read and checked for completion during next runs.

7. Litrepl provides \textbf{start}, \textbf{stop}, \textbf{restart} commands
accepting the name of interpreter class to act on to start, stop or restart
interpreter sessions.

----

%lnoai
\end{comment}

%lresult
Litrepl's ability to maintain interpreter sessions in the background is crucial
for enabling a Read-Eval-Print Loop (REPL) environment. These resources, shown
in Figure \ref{fig:resource-allocation}, are stored as files within an auxiliary
directory.

If not specified by command-line arguments or environment variables, the
directory path is automatically derived from the interpreter class name, the
current working directory, and the OS's temporary file location.

The auxiliary directory includes two POSIX pipes for interpreter I/O and a file
recording the running interpreter's process ID, aiding session management.

When a code section is evaluated, Litrepl assigns a response file name derived
from hashing the code. This response file stores the output from the
interpreter.

During evaluation, Litrepl spawns a response reader process with a soft lock,
active until the interpreter completes and responds to an echo probe. If the
response exceeds the configured duration, Litrepl outputs a partial result tag,
which is recognized and reevaluated in subsequent runs. Figure
\ref{fig:partial-result} shows an example partial result section.

Litrepl provides \textbf{start}, \textbf{stop}, and \textbf{restart} commands to
control background sessions. The \textbf{interrupt} command sends an
interruption signal to the interpreter. Finally, the \textbf{repl} command
establishes direct communication with the interpreter, allowing manual
inspection of its state.
%lnoresult

@@

%lignore
\begin{figure*}[hbt!]
  \centering
  \begin{minipage}{\textwidth}
  \begin{verbatim}
  document       ::= (code | result | ignore | text)*
  code           ::= (code-normal | code-comment)
  result         ::= (result-normal | result-comment)
  code-normal    ::= "\begin{MARKER}" text "\end{MARKER}"
  code-comment   ::= "% MARKER" text "% noMARKER"
  result-normal  ::= "\begin{result}" text "\end{result}"
  result-comment ::= "% result" text "% noresult"
  ignore         ::= "% ignore" text "% noignore"
  text           ::= ...
  \end{verbatim}
  \end{minipage}
  \caption{An illustrative grammar template for LaTeX documents where marker
  serves as a parameter configured via command-line arguments for each supported
  interpreter class.}
  \label{fig:document_representation}
\end{figure*}
%lnoignore
